---
output: pdf_document
# 112077423
---

## Student ID: 112077423

```{r message=FALSE}
library(ggplot2)
```

```{r}
# read file
customers_df <- read.delim('customers.txt')
# print first 6 elements in dataframe
head(customers_df)
```

## Question 1

*Find the 5th element in the original list.*

```{r}
fifth <- customers_df[5,]
print(paste('The 5th element is', fifth, sep=' '))
```

## Question 2

*Find the fifth lowest age.*

```{r}
# sort in ascending order
tmp <- customers_df[order(customers_df$age, decreasing=FALSE), ]
fifth_lowest <- tmp[5]
print(paste('The 5th lowest element is', fifth_lowest, sep=' '))
```

## Question 3

*Extract the five lowest ages together.*

```{r}
five_lowest <- head(tmp, 5)
five_lowest
```

\newpage
## Student ID: 112077423

## Question 4

*Get the five highest ages by first sorting them in decreasing order first.*

```{r}
# sort in descending order
tmp2 <- customers_df[order(customers_df$age, decreasing=TRUE), ]
five_highest <- head(tmp2, 5)
five_highest
```

## Question 5

*Find the average (mean) age.*

```{r}
mean_age <- mean(customers_df$age)
print(paste('The average is', mean_age, sep=' '))
```

## Question 6

*Find the standard deviation of ages.*

```{r}
std <- sd(customers_df$age)
print(paste('The standard deviation is', std, sep=' '))
```

## Question 7

*Create a new variable with the difference between each age and the mean age.*

```{r}
age_diff <- customers_df$age - mean_age
```

## Question 8

*Find the average of age_diff.*

```{r}
diff_mean <- mean(age_diff)
diff_mean
```

\newpage
## Student ID: 112077423

## Question 9a

*Visualize the raw data in form of histogram.*

```{r fig.align='center'}
ggplot(data=customers_df, aes(x=age)) + 
  geom_histogram(color="black", fill="white", bins=20) +
  theme_minimal() 
```

\newpage
## Student ID: 112077423

## Question 9b

*Visualize the raw data in form of density plot.*

```{r fig.align='center'}
ggplot(data=customers_df, aes(x=age)) + 
  geom_density(color="darkblue") +
  theme_minimal() 
```

\newpage
## Student ID: 112077423

## Question 9c

*Visualize the raw data in form of boxplot+stripchart.*

```{r fig.align='center'}
ggplot(data=customers_df, aes(x='', y = age)) + 
  geom_boxplot() +
  geom_jitter(position=position_jitter(0.2), color='darkblue', shape=17) + 
  theme_minimal() +
  labs(x="")
```
