---
output:
  pdf_document:
    latex_engine: xelatex
---

## Student ID: 112077423

```{r message=FALSE}
library(ggplot2)
library(compstatslib)
library(data.table)
library(tidyr)
library(lsa)
```

```{r}
cars <- read.table("auto-data.txt", header=FALSE, na.strings = "?")

names(cars) <- c("mpg", "cylinders", "displacement", "horsepower", "weight", 
                 "acceleration", "model_year", "origin", "car_name")

cars_log <- with(cars, data.frame(log(mpg), log(weight), log(acceleration), 
                                  model_year, origin))
head(cars_log)
```

## Question 1(a)

```{r}
mean_wt <- log(mean(cars$weight))
lw <- cars_log[cars_log$log.weight. < mean_wt,]
hw <- cars_log[cars_log$log.weight. > mean_wt,]

lw_regr <- lm(log.mpg. ~ log.acceleration., data=lw)
hw_regr <- lm(log.mpg. ~ log.acceleration., data=hw)

ggplot() +
  geom_point(data = lw, aes(x=log.acceleration., y=log.mpg., color='light')) +
  geom_point(data = hw, aes(x=log.acceleration., y=log.mpg., color='heavy')) +
  geom_abline(slope=lw_regr$coefficients[2], intercept=lw_regr$coefficients[1], color="cyan4") +
  geom_abline(slope=hw_regr$coefficients[2], intercept=hw_regr$coefficients[1], color="coral") +
  labs(title='acceleration vs. mpg', color='Cars') +
  theme_minimal()
```

## Question 1(b)

```{r}
full_lw_regr <- lm(log.mpg. ~ . - origin + factor(origin), data=lw) 
summary(full_lw_regr)
```

The model is significant at alpha=0.01. Adjusted R-squared is 0.72, meaning that ~72% of variation in the dependent variable can be explained by the variation in the independent variables. All the variables are significant at alpha=1% except log.acceleration. and origin.

```{r}
full_hw_regr <- lm(log.mpg. ~ . - origin + factor(origin), data=hw) 
summary(full_hw_regr)
```

The model is significant at alpha=0.01. Adjusted R-squared is 0.71, meaning that ~71% of variation in the dependent variable can be explained by the variation in the independent variables. All the variables are significant at alpha=1% except log.acceleration. and origin.

## Question 1(c)

- For heavy cars, there are only two countries present;

- mpg is higher for light cars;

- Looking at the scatter plot, heavy cars' mpg rises more as acceleration increases.

## Question 2(a)

I think weight is a moderator and acceleration is an independent variable.

## Question 2(b)

```{r}
model1 <- lm(log.mpg. ~ . - origin + factor(origin), data=cars_log)
summary(model1)
```

```{r}
model2 <- lm(log.mpg. ~ log.weight. + log.acceleration. + 
             log.weight. * log.acceleration. + 
             model_year + factor(origin), data=cars_log)
summary(model2)
```

```{r}
acl_mc <- scale(cars_log$log.acceleration., center=TRUE, scale=FALSE)
wt_mc <- scale(cars_log$log.weight., center=TRUE, scale=FALSE)

model3 <- lm(cars_log$log.mpg. ~ wt_mc + acl_mc +
             wt_mc * acl_mc + 
             cars_log$model_year + factor(cars_log$origin))
summary(model3)
```

```{r}
wt_x_acl <- cars_log$log.weight. * cars_log$log.acceleration.
interaction_regr <- lm(wt_x_acl ~ cars_log$log.weight. + cars_log$log.acceleration.)
interaction_ortho <- interaction_regr$residuals

model4 <- lm(cars_log$log.mpg. ~ cars_log$log.weight. + cars_log$log.acceleration. + 
             cars_log$model_year + factor(cars_log$origin) + interaction_ortho)
summary(model4)
```

## Question 2(c)

```{r}
# raw
cor(cbind(log.mpg = cars_log$log.mpg.,
          log.weight.=cars_log$log.weight., log.acceleration.=cars_log$log.acceleration.,
          intxn = cars_log$log.weight. * cars_log$log.acceleration.))
```
- The correlation between that interaction term and log.weight. is 0.108

- The correlation between that interaction term and log.acceleration. is 0.852

```{r}
# mean-centered
tmp <- cor(cbind(log.mpg = cars_log$log.mpg.,
                 wt_mc=wt_mc, acl_mc=acl_mc, intxn = wt_mc * acl_mc))
colnames(tmp) <- c('log.mpg', 'wt_mc', 'acl_mc', 'intxn')
rownames(tmp) <- c('log.mpg', 'wt_mc', 'acl_mc', 'intxn')
tmp
```
 
- The correlation between that interaction term and mean-centered weight is -0.2

- The correlation between that interaction term and mean-centered accelaration is 0.35

```{r}
# orthogonalized
round(cor(cbind(log.mpg = cars_log$log.mpg.,
                log.weight.=cars_log$log.weight.,
                log.acceleration.=cars_log$log.acceleration., interaction_ortho)), 2)
```

- The correlation between that interaction term and log.weight. is 0

- The correlation between that interaction term and log.acceleration. is 0

## Question 3(a)

```{r}
cars_log <- with(cars, data.frame(log(mpg), log(weight), log(acceleration),
                                  log(cylinders), model_year, origin))
head(cars_log)
```

```{r}
model1 <- lm(log.weight. ~ log.cylinders., data=cars_log)
summary(model1)
```

log.cylinders. is significant at alpha=0.01.

```{r}
model2 <- lm(log.mpg. ~ log.weight. + log.acceleration. + 
             model_year + factor(origin), data=cars_log)
summary(model2)
```

The weight has a significant direct effect on mpg at alpha=0.01. 

## Question 3(b)

```{r}
tmp <- model1$coefficients[2] * model2$coefficients[2]
cat('indirect effect of cylinders on mpg =', tmp, sep=' ')
```
## Question 3(c)

```{r}
boot_mediation <- function(model1, model2, dataset) {
  boot_index <- sample(1:nrow(dataset), replace=TRUE)
  data_boot <- dataset[boot_index, ]
  regr1 <- lm(model1, data_boot)
  regr2 <- lm(model2, data_boot)
  return(regr1$coefficients[2] * regr2$coefficients[2])
}

set.seed(645218)

indirect <- replicate(2000, boot_mediation(model1, model2, cars_log))
quantile(indirect, probs=c(0.025, 0.975))
```

```{r}
plot(density(indirect), main='distribution of the indirect effect', lwd=1.5)
abline(v=quantile(indirect, probs=c(0.025, 0.975)), lw=1.5, lty='dashed', col='red')
```